---
title: "DSD data analysis"
author: "Nelly"
date: "2/17/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(pacman)
pacman::p_load(tidyverse, # Manipulate data
               sjPlot, # make chi-square test table
               gmodels, # mosaic plot
               lubridate) # Date manipulation


library(dplyr)
library(blorr)
library(aod)
library(ggplot2)
library(readxl)
library(stringr)
library(scales)
library(reprex)
library(ggcorrplot)
library(rcompanion)

```
## 


 Load Data 
```{r}
X2_DSD_Final_Analysis <- read_excel("2-DSD Final Analysis.xlsx")

df<-X2_DSD_Final_Analysis

```
Global Variables 
```{r}
curr_year <- 2022
```





 ## data cleaning 
```{r}
#spliting organization
df <-df %>% 
  tidyr::separate(Organisation, sep = " / ", into = c("Region", "SNU", "PSNU", "Facility"))
table(df$Region)
```

 
 
 #Recode variables- Gender 
```{r}

df$Gender = as.factor(df$Gender)
df$Gender = recode(df$Gender,
                          '1' = "Male",
                          '2' = "Female"
)     
```
 
  
```{r}
df$TXDuration    = as.factor(df$TXDuration)
df$TXDuration = recode(df$TXDuration,
                          '1' = "Less than 3 months ",
                          '2' = "Between 3 and 6 months",
                          '3' = "Between 6 months and a Year",
                          '4' = "More than an Year"
)   
```
 
```{r}
df$VLC = as.factor(df$VLC)
df$VLC = recode(df$VLC,
                          '1' = "Yes ",
                          '2' = "No"
)  
```

 
```{r}
df$VLS = as.factor(df$VLS)
df$VLS = recode(df$VLS,
                          '1' = "Yes ",
                          '2' = "No"
)  
```
 

```{r}
df$IIT = as.factor(df$IIT)
df$IIT = recode(df$IIT,
                          '1' = "Yes ",
                          '2' = "No"
)  
```
 

 
```{r}
df$Person_administered = as.factor(df$Person_administered)
df$Person_administered = recode(df$Person_administered,
                          '1' = "Client",
                          '2' = "Caretaker"
) 
```
 
```{r}
df$MMD_last_visit = as.factor(df$MMD_last_visit)
df$MMD_last_visit = recode(df$MMD_last_visit,
                          '1' = "Less than one month if ART ",
                          '2' = "1-2 Months",
                          '3' = "3-5 months ",
                          '4' = "6+ months"
) 
```
 
 
```{r}
df$preferred_MMD= as.factor(df$preferred_MMD)
df$preferred_MMD = recode(df$preferred_MMD,
                          '1' = "Less than one month if ART ",
                          '2' = "1-2 Months",
                          '3' = "3-5 months ",
                          '4' = "6+ months"
) 
```


```{r}
df$Current_ARV_Deivery_Model = as.factor(df$Current_ARV_Deivery_Model)
df$Current_ARV_Deivery_Model = recode(df$Current_ARV_Deivery_Model,
                          '1' = "Facility Based individual management",
                          '2' = "Facility based group ",
                          '3' = "facility track refill ",
                          '4' = "Community Led ART Delivery ",
                          '5' = "community Pharmacy",
                          '6' =" other "
) 
```


```{r}
df$ISCurrent_PrefDSD = as.factor(df$ISCurrent_PrefDSD)
df$ISCurrent_PrefDSD = recode(df$ISCurrent_PrefDSD,
                          '1' = "Yes ",
                          '2' = "No"
)
```
 
 
```{r}

df$Preferred_DSD = as.factor(df$Preferred_DSD)
df$Preferred_DSD = recode(df$Preferred_DSD,
                          '1' = "Facility Based individual management",
                          '2' = "Facility based group ",
                          '3' = "Facility track refill ",
                          '4' = "Community Led ART Delivery ",
                          '5' = "community drug distribution",
                          '6' = "Community Pharmacy ",
                          '7' = "Other"
) 

```
 
```{r}
df$CHW_Present = as.factor(df$CHW_Present)
df$CHW_Present = recode(df$CHW_Present,
                          '1' = "Yes ",
                          '2' = "No"
)  
```
 
```{r}
df$CHW_Follow_Up = as.factor(df$CHW_Follow_Up)
df$CHW_Follow_Up = recode(df$CHW_Follow_Up,
                          '1' = "Yes ",
                          '2' = "No"
)  
```


```{r}
df$`Contacted for follow-up` = as.factor(df$`Contacted for follow-up`)
df$`Contacted for follow-up` = recode(df$`Contacted for follow-up`,
                          '1' = "Yes ",
                          '2' = "No"
)  
```
#diving the data by peds and adults 
## use age bands instead 
  CLient age to peds and Adults

```{r}
df <- df %>%  
  mutate(agecoarse = if_else(Current_CLIENTAGE  < 15, "<15", "15+"))
table(df$agecoarse)
```




```{r}
df$Current_CLIENTAGE = cut(df$Current_CLIENTAGE,
                   breaks = c(0,4,9,14,19,24,29,34,39,44,49,100),
                   labels = c("0-04","05-09","10-14","15-19", "20-24", "25-29", "30-34", "35-39","40-44", "45-49","50+"))
table(df$Current_CLIENTAGE)

```






#   creating categories out of  HIV Diagnosis Year 

```{r}

#str(df$HIV_DIagnosisYear)
df <- df %>% 
  mutate(TimewHIV= 2022-(HIV_DIagnosisYear))

#curr_year <- 2022( assign Global variable)
df$TimewHIV = cut(df$TimewHIV,
                   breaks = c(0,10,20,40),
                   labels = c("01-09","10-19","20+"))
table(df$TimewHIV)

```

#distribution by age and gender 

```{r}
#Gender_age <- xtabs(~Current_CLIENTAGE+Gender, data = df) 
#view(Gender_age)
```
#Frequencies 
# visualizing distributions 
```{r}
ggplot(data = df) +
  geom_bar(mapping = aes(x = Current_CLIENTAGE))
df
```


```{r}
ggplot(data = df) +
  geom_bar(mapping = aes(x = agecoarse))

```




```{r}
ggplot(data = df) +
  geom_bar(mapping = aes(x = Gender))

```


```{r}
ggplot(data = df) +
  geom_histogram(mapping = aes(x = HIV_DIagnosisYear), binwidth = 0.5)

```

# how many patients are recieving ARVs using their most prefered Method of dispensation 
```{r}

ggplot(data = df) +
  geom_bar(mapping = aes(x = Preferred_DSD))

```
```{r}
ggplot(data = df) +
  geom_bar(mapping = aes(x =ISCurrent_PrefDSD))
```





```{r}
ggplot(data = df) +
  geom_bar(mapping = aes(x = IIT))
```

```{r}
ggplot(data = df) +
  geom_bar(mapping = aes(x = MMD_last_visit))
```

```{r}
mosaicplot(preferred_MMD ~ MMD_last_visit ,data = df)

```


```{r}
mosaicplot(Region ~ IIT ,data = df)
```

```{r}
mosaicplot(agecoarse ~ IIT ,data = df)
```



```{r}
mosaicplot(agecoarse ~ VLS ,data = df)
```




```{r}
mosaicplot(Preferred_DSD ~ IIT ,data = df)
```



```{r}
mosaicplot(Preferred_DSD ~ Current_ARV_Deivery_Model ,data = df)
```

#  Reg



#Statistical analysis 

```{r}
df %>%
  select(Current_CLIENTAGE, VLS) %>% 
  sjtab(fun = "xtab", var.labels=c("Current_CLIENTAGE", "VLS"),
        show.row.prc=T, show.col.prc=T, show.summary=T, show.exp=T, show.legend=T)
```

 
 
 
```{r}
df %>%
  select(Gender, VLS) %>% 
  sjtab(fun = "xtab", var.labels=c("Gender", "VLS"),
        show.row.prc=T, show.col.prc=T, show.summary=T, show.exp=T, show.legend=T)

```


```{r}
df %>%
  select(Region, VLS) %>% 
  sjtab(fun = "xtab", var.labels=c("Region", "VLS"),
        show.row.prc=T, show.col.prc=T, show.summary=T, show.exp=T, show.legend=T)
```


```{r}
df %>%
  select(IIT, VLS) %>% 
  sjtab(fun = "xtab", var.labels=c("IIT", "VLS"),
        show.row.prc=T, show.col.prc=T, show.summary=T, show.exp=T, show.legend=T)
```



```{r}
df %>%
  select(CHW_Present, VLS) %>% 
  sjtab(fun = "xtab", var.labels=c("CHW_Present", "VLS"),
        show.row.prc=T, show.col.prc=T, show.summary=T, show.exp=T, show.legend=T)
```

```{r}

df %>%
  select(TXDuration, VLS) %>% 
  sjtab(fun = "xtab", var.labels=c("TXDuration", "VLS"),
        show.row.prc=T, show.col.prc=T, show.summary=T, show.exp=T, show.legend=T)

```

```{r}
df %>%
  select(VLC, VLS) %>% 
  sjtab(fun = "xtab", var.labels=c("VLC", "VLS"),
        show.row.prc=T, show.col.prc=T, show.summary=T, show.exp=T, show.legend=T)
```



```{r}
df %>%
  select(MMD_last_visit, VLS) %>% 
  sjtab(fun = "xtab", var.labels=c("MMD_last_visit", "VLS"),
        show.row.prc=T, show.col.prc=T, show.summary=T, show.exp=T, show.legend=T)
```

```{r}
df %>%
  select(TimewHIV, VLS) %>% 
  sjtab(fun = "xtab", var.labels=c("TimewHIV", "VLS"),
        show.row.prc=T, show.col.prc=T, show.summary=T, show.exp=T, show.legend=T)
```



```{r}
logit.fit_1 = glm(VLS ~ Current_CLIENTAGE + Gender + Region + IIT + TXDuration + MMD_last_visit + TimewHIV,  data = df, family = "binomial")
summary(logit.fit_1)
```

```{r}
cbind(exp(coefficients(logit.fit_1)), exp(confint(logit.fit_1)))
```


 VLC
 
```{r}
df %>%
  select(MMD_last_visit,VLC ) %>% 
  sjtab(fun = "xtab", var.labels=c("MMD_last_visit", "VLC"),
        show.row.prc=T, show.col.prc=T, show.summary=T, show.exp=T, show.legend=T)
```

 
```{r}
df %>%
  select(Gender,VLC ) %>% 
  sjtab(fun = "xtab", var.labels=c("Gender", "VLC"),
        show.row.prc=T, show.col.prc=T, show.summary=T, show.exp=T, show.legend=T)
```
 
 
 
```{r}
df %>%
  select(Current_CLIENTAGE,VLC ) %>% 
  sjtab(fun = "xtab", var.labels=c("Current_CLIENTAGE", "VLC"),
        show.row.prc=T, show.col.prc=T, show.summary=T, show.exp=T, show.legend=T)
```
 
```{r}
df %>%
  select(Region,VLC ) %>% 
  sjtab(fun = "xtab", var.labels=c("Region", "VLC"),
        show.row.prc=T, show.col.prc=T, show.summary=T, show.exp=T, show.legend=T)
```
 

```{r}
df %>%
  select(IIT,VLC ) %>% 
  sjtab(fun = "xtab", var.labels=c("IIT", "VLC"),
        show.row.prc=T, show.col.prc=T, show.summary=T, show.exp=T, show.legend=T)
```


```{r}
df %>%
  select(CHW_Present,VLC ) %>% 
  sjtab(fun = "xtab", var.labels=c("CHW_Present", "VLC"),
        show.row.prc=T, show.col.prc=T, show.summary=T, show.exp=T, show.legend=T)
```
 
 
```{r}
df %>%
  select(TXDuration,VLC ) %>% 
  sjtab(fun = "xtab", var.labels=c("TXDuration", "VLC"),
        show.row.prc=T, show.col.prc=T, show.summary=T, show.exp=T, show.legend=T)
```
 
 
```{r}
df %>%
  select(VLS,VLC ) %>% 
  sjtab(fun = "xtab", var.labels=c("VLS", "VLC"),
        show.row.prc=T, show.col.prc=T, show.summary=T, show.exp=T, show.legend=T)
```
 
```{r}
df %>%
  select(TimewHIV,VLC ) %>% 
  sjtab(fun = "xtab", var.labels=c("TimewHIV", "VLC"),
        show.row.prc=T, show.col.prc=T, show.summary=T, show.exp=T, show.legend=T)
```



```{r}
logit.fit_2 = glm(VLC ~ Current_CLIENTAGE + VLS + Region + IIT + TXDuration + MMD_last_visit + TimewHIV,  data = df, family = "binomial")
summary(logit.fit_2)
```


```{r}
cbind(exp(coefficients(logit.fit_2)), exp(confint(logit.fit_2)))
```
 
  CHW Present 
  
```{r}
df %>%
  select(MMD_last_visit,CHW_Present ) %>% 
  sjtab(fun = "xtab", var.labels=c("MMD_last_visit", "CHW_Present"),
        show.row.prc=T, show.col.prc=T, show.summary=T, show.exp=T, show.legend=T)
```
 
 


```{r}
df %>%
  select(Region,CHW_Present ) %>% 
  sjtab(fun = "xtab", var.labels=c("Region", "CHW_Present"),
        show.row.prc=T, show.col.prc=T, show.summary=T, show.exp=T, show.legend=T)
```
  
  
```{r}
df %>%
  select(Gender,CHW_Present ) %>% 
  sjtab(fun = "xtab", var.labels=c("Gender", "CHW_Present"),
        show.row.prc=T, show.col.prc=T, show.summary=T, show.exp=T, show.legend=T)
```
  
  
```{r}
df %>%
  select(IIT,CHW_Present ) %>% 
  sjtab(fun = "xtab", var.labels=c("IIT", "CHW_Present"),
        show.row.prc=T, show.col.prc=T, show.summary=T, show.exp=T, show.legend=T)
```


```{r}
df %>%
  select(TXDuration,CHW_Present ) %>% 
  sjtab(fun = "xtab", var.labels=c("TXDuration", "CHW_Present"),
        show.row.prc=T, show.col.prc=T, show.summary=T, show.exp=T, show.legend=T)
```


```{r}
df %>%
  select(Current_CLIENTAGE,CHW_Present ) %>% 
  sjtab(fun = "xtab", var.labels=c("Current_CLIENTAGE", "CHW_Present"),
        show.row.prc=T, show.col.prc=T, show.summary=T, show.exp=T, show.legend=T)
```


```{r}
df %>%
  select(Current_ARV_Deivery_Model,CHW_Present ) %>% 
  sjtab(fun = "xtab", var.labels=c("Current_ARV_Deivery_Model", "CHW_Present"),
        show.row.prc=T, show.col.prc=T, show.summary=T, show.exp=T, show.legend=T)

```
  
```{r}
df %>%
  select(TimewHIV,CHW_Present ) %>% 
  sjtab(fun = "xtab", var.labels=c("TimewHIV", "CHW_Present"),
        show.row.prc=T, show.col.prc=T, show.summary=T, show.exp=T, show.legend=T)
```
  
  
  
```{r}
logit.fit_3 = glm(VLC ~ Current_CLIENTAGE + VLS + Region + IIT + TXDuration + MMD_last_visit + Current_ARV_Deivery_Model + TimewHIV,  data = df, family = "binomial")
summary(logit.fit_3)
```
  
```{r}
cbind(exp(coefficients(logit.fit_3)), exp(confint(logit.fit_3)))
```
  



```{r}
df %>%
  select(Current_CLIENTAGE, IIT) %>% 
  sjtab(fun = "xtab", var.labels=c("Current_CLIENTAGE", "IIT"),
        show.row.prc=T, show.col.prc=T, show.summary=T, show.exp=T, show.legend=T)
```


```{r}
df %>%
  select(Gender, IIT) %>% 
  sjtab(fun = "xtab", var.labels=c("Gender", "IIT"),
        show.row.prc=T, show.col.prc=T, show.summary=T, show.exp=T, show.legend=T)
```
  
  
  
```{r}
df %>%
  select(TXDuration, IIT) %>% 
  sjtab(fun = "xtab", var.labels=c("TXDuration", "IIT"),
        show.row.prc=T, show.col.prc=T, show.summary=T, show.exp=T, show.legend=T)
```
  
  
  
```{r}
df %>%
  select(VLS, IIT) %>% 
  sjtab(fun = "xtab", var.labels=c("VLS", "IIT"),
        show.row.prc=T, show.col.prc=T, show.summary=T, show.exp=T, show.legend=T)
```
 
 

```{r}
df %>%
  select(MMD_last_visit, IIT) %>% 
  sjtab(fun = "xtab", var.labels=c("MMD_last_visit", "IIT"),
        show.row.prc=T, show.col.prc=T, show.summary=T, show.exp=T, show.legend=T)
```
  
  
```{r}
df %>%
  select(CHW_Present, IIT) %>% 
  sjtab(fun = "xtab", var.labels=c("CHW_Present", "IIT"),
        show.row.prc=T, show.col.prc=T, show.summary=T, show.exp=T, show.legend=T)
```
  
```{r}
df %>%
  select(Current_ARV_Deivery_Model, IIT) %>% 
  sjtab(fun = "xtab", var.labels=c("Current_ARV_Deivery_Model", "IIT"),
        show.row.prc=T, show.col.prc=T, show.summary=T, show.exp=T, show.legend=T)
  
```
  
```{r}
df %>%
  select(TimewHIV,IIT) %>% 
  sjtab(fun = "xtab", var.labels=c("TimewHIV", "IIT"),
        show.row.prc=T, show.col.prc=T, show.summary=T, show.exp=T, show.legend=T)
```
  
```{r}
df %>%
  select(Region, IIT) %>% 
  sjtab(fun = "xtab", var.labels=c("Region", "IIT"),
        show.row.prc=T, show.col.prc=T, show.summary=T, show.exp=T, show.legend=T)
```
  

```{r}
df %>%
  select(Preferred_DSD, IIT) %>% 
  sjtab(fun = "xtab", var.labels=c("Preferred_DSD", "IIT"),
        show.row.prc=T, show.col.prc=T, show.summary=T, show.exp=T, show.legend=T)

```

df
  
```{r}
logit.fit_4 = glm(IIT ~ Current_CLIENTAGE + VLS + Region +VLC + TXDuration + MMD_last_visit + Current_ARV_Deivery_Model+ TimewHIV,  data = df, family = "binomial")
summary(logit.fit_4)
```
```{r}
cbind(exp(coefficients(logit.fit_4)), exp(confint(logit.fit_4)))
```
Corrplot 
```{r}
model.matrix(~ 0+., data=df) %>% 
  cor(use="pairwise.complete.obs") %>% 
  ggcorrplot(show.diag = F, type="lower", lab=TRUE, lab_size=2)
```


```{r}
mixed_assoc = function(df, cor_method="spearman", adjust_cramersv_bias=TRUE){
    df_comb = expand.grid(names(df), names(df),  stringsAsFactors = F) %>% set_names("X1", "X2")

    is_nominal = function(x) class(x) %in% c("factor", "character")
    # https://community.rstudio.com/t/why-is-purr-is-numeric-deprecated/3559
    # https://github.com/r-lib/rlang/issues/781
    is_numeric <- function(x) { is.integer(x) || is_double(x)}

    f = function(xName,yName) {
        x =  pull(df, xName)
        y =  pull(df, yName)

        result = if(is_nominal(x) && is_nominal(y)){
            # use bias corrected cramersV as described in https://rdrr.io/cran/rcompanion/man/cramerV.html
            cv = cramerV(as.character(x), as.character(y), bias.correct = adjust_cramersv_bias)
            data.frame(xName, yName, assoc=cv, type="cramersV")

        }else if(is_numeric(x) && is_numeric(y)){
            correlation = cor(x, y, method=cor_method, use="complete.obs")
            data.frame(xName, yName, assoc=correlation, type="correlation")

        }else if(is_numeric(x) && is_nominal(y)){
            # from https://stats.stackexchange.com/questions/119835/correlation-between-a-nominal-iv-and-a-continuous-dv-variable/124618#124618
            r_squared = summary(lm(x ~ y))$r.squared
            data.frame(xName, yName, assoc=sqrt(r_squared), type="anova")

        }else if(is_nominal(x) && is_numeric(y)){
            r_squared = summary(lm(y ~x))$r.squared
            data.frame(xName, yName, assoc=sqrt(r_squared), type="anova")

        }else {
            warning(paste("unmatched column type combination: ", class(x), class(y)))
        }

        # finally add complete obs number and ratio to table
        result %>% mutate(complete_obs_pairs=sum(!is.na(x) & !is.na(y)), complete_obs_ratio=complete_obs_pairs/length(x)) %>% rename(x=xName, y=yName)
    }

    # apply function to each variable combination
    map2_df(df_comb$X1, df_comb$X2, f)
}
```


```{r}
mixed_assoc(df)
```
```{r}
view(mixed_assoc)

```




 